### 2021-02-11 paper/paper.tex
added lots of equations to algorithm section

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Designing and Developing a Personalised Recommender System}

\author{\IEEEauthorblockN{lrfk99}
\IEEEauthorblockA{\textit{dept. Computer Science} \\
\textit{Durham University}\\
Durham, United Kingdom}
}

\maketitle

\section{Introduction}

\subsection{Domain of application}
The domain of this hybrid recommender system is restaurants, with a focus on trying new restaurants. 
This system is built around restaurants in a particular metropolitan area which serve a particular 
type of food.

\subsection{Related work review}
Hybrid recommender systems which combine knowledge-based (KB) and collaborative-filtering (CF) techniques 
are designed to combine the ability of KB techniques to give generic recommendations to new users with 
small profiles and the strength of CF to find peer users with unexpected shared preferences. 
This combination of techniques is particularly well suited to restaurants since a user may often be in the 
mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
type or restaurant, which is where the knowledge-based recommender is helpful. 
A collaborative-filtering algorithm will then be able to order all the restaurants which match this 
explicit criteria by how the user should rate them, based on the opinions of other similar users, 
helping them to discover new restaurants to try. 
These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
In particular, Burke's 1999 paper \cite{burke1999integrating} is the first to discuss specifically using these 
two techniques in the same cascaded style that I implement here.

\subsection{Purpose/Aim}
The purpose of this application is to give suitable suggestions of restaurants for a user to go to. 
The recommended restaurants should be in the metropolitan area which the user specifies, and serve the 
type of food that the user asks for. 
As well as fitting this explicit criteria, the suggested restaurants should be similar to other restaurants 
that the user has rated highly in the past and which other users with similar preferences also like. 
The system will allow the user to add ratings of the restaurants they are recommended, and these ratings will 
affect future suggestions the system gives for that user and other similar users.

\section{Methods}

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
1 in the covid section. 
The $business.json$ file contains business data including location data, attributes, and categories. 
$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
that the review is written for. 
The file $user.json$ holds every user's first name along with other data adn metadata which I don't use \cite{YelpDataset}. 
The data files $checkin.json$ and $tip.json$ are not used in my system. 
The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
the interface is command line only, so they were discarded. 
The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
special message for customers. 
Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
preparation is required to save computation time and in order to provide more accurate predictions. 
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
covid dataset removed. 
The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).

\subsection{Hybrid scheme}
The two personalised recommender systems I implement are knowledge-based and collaborative-filtering, and 
the hybrid scheme is cascade. 
A cascade hybrid scheme works particularly well in my domain of restaurants with an emphasis on providing accurate 
predictions of restaurants which a user may like in a city they have never been to before, based on the restaurants 
they like in their home-city. 
This is because the knowledge-based system can filter the restaurants by location and by the type of food they serve, 
while the collaborative-filtering system is able to find restaurants which are similar to the ones the user has 
previously rated highly. 
So when these two systems are combined in a cascade style with the knowledge-based system as the primary recommender 
and the collaborative-filtering system as the secondary the hybrid system can find hidden similarities between 
restaurants in different cities very accurately. 
This method of hybrid recommender system (HRS) has been implemented as a restaurant recommender in the Entree 
system, giving very accurate results \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}.

\subsection{Recommendation techniques/algorithms}
The first prediction method used in this hybrid system is a knowledge-based (KB) approach, which relies on knowledge 
about each of the restaurants based on its location and the type of food that it serves. 
This approach completely eliminates the cold-start problem, a new user will still receive relevant recommendations 
for the city they are in and the type of food they in the mood for when using a KB system. 
The second prediction method is a collaborative-filtering (CF) approach, specifically Single Value Decomposition (SVD) 
matrix factorisation. 
This approach was chosen because it scales well with large datasets, which is important since my dataset has 
~120,000 entries. 
The CF technique here also introduces an element of variability in the recommendations, such that the user 
may not receive the same ordering of suggestions when they run the system multiple times with the same 
KB specifications, helping the user to try new things. 
This system allows the user to make ratings through the user interface, providing the ability to experiment with 
the system in terms of how new ratings affect the current user's recommendations as well as those of other similar users.

\subsection{Evaluation methods}
How to evaluate

\section{Implementation}

\subsection{Input interface}
The system has a command line based interface, so all user input of the input is given via the command line. 
The system recognises the active user by their unique user ID, some example IDs are provided in the 'README.txt' 
file for testing and demonstration purposes. 
Only explicit user data is gathered in order to make the system more explainable and transparent, 
and users are made aware of which data is collected, how it is collected, and for which purposes. 
This data includes the user's nearest city (of the 10 available) and the type of food that 
they are in the mood for. 
The user is able to change these choices from the main menu, as well as switch to a different user 
for testing purposes. 
When the user is presented with a list of recommendations, they can then choose one that they want 
to see more information about and from here they can leave a rating, return to the list, or return to 
the main menu. 
User profiles can be updated by giving a rating of a restaurant through the system. 
Any rating given will affect future recommendations given by the system to that user 
and to other similar users.

\subsection{Recommendation algorithm}
The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
If a user has not rated a restaurant then the rating in that cell is set to 0. 
The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
\begin{equation}
    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
\end{equation}
Where the parameters are learnt by minimising the following regularised squared error: 
\begin{equation}
    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
\end{equation}
This minimisation is performed by stochastic gradient descent:
\begin{align}
    \begin{split}
    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
    \end{split}
\end{align}
where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}.

The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
I used a model-based approach to collaborative-filtering based on matrix factorisation. 
The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
and in the specified city, I predict what the user would rate that restaurant. 
Then I sort those restaurants by their score and take the top 8 to present to the user.

\subsection{Output interface}
Recommendations and prediction scores are presented to the user via the command line interface and using an ASCII table. 
I used a python module called texttable for this, since it makes the output very clear and visually appealing. 
I chose to output the top 8 recommendations since I think this is enough variety and choice for the user to find 
a restaurant to try, without there being too many options which might overwhelm the user. 
If I presented more than 8 recommendations at a time, the table wouldn't fit in a single terminal output window and 
would require the user to scroll around to read different parts of the output. 
Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
an ASCII table is the most visually attractive way to present information, and the white text on a (typically) black 
background will help accommodate users who are colourblind or who require a high-contrast interface. 
The interface is clear enough to understand that users of all ages are able to use the system, and if a user 
requires the text to be larger then they can increase the text size of their terminal. 
Along with the recommended restaurants, an explanation of how these results are calculated is also provided, as 
well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
This, along with an option in the main menu to view information about the data the system uses and how it is used, 
further adds to the explainability and transparency. 
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{output.png}
    \caption{System output}
    \label{fig:output}
\end{figure}

\section{Evaluation results}

\subsection{Comparison against baseline implementation}
Compare vs generic suggestions

\subsection{Comparison against hybrid recommenders in related studies}
Read some papers

\subsection{Ethical issues}
People's personal data

\section{Conclusion}

\subsection{Limitations}
What can it not do?

\subsection{Further developments}
What could I do in the future \cite{5284958}. 
Implement the idea of particular categories being closer some than others. 
Specify price range and/or consider price when comparing similarities of restaurants. 
Include demographic information

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}

\end{document}